Sufficient Conditions for Reachability in Automata Networks with Priorities

### 2015-02-13 parts/example.tex
detailed small example with scenario extraction

Finally, the static analysis results presented in
\pref{ssec:ua} and \pref{ssec:ordered-ua} can be used to check if the model is functional,
\ie if gene $a$ can oscillate, thus leading to the formation of stripes.
Starting from context $\ctx = \PHstate{f_1, a_0, c_0}$,
we thus want to check the reachability of
$\w = \PHobj{a_0}{a_1} \PHconcat \PHobj{a_1}{a_0} \PHconcat \PHobj{a_0}{a_1}$.
Rather than checking this whole objective sequence with \pref{th:approxinf},
we break it using \pref{thm:ordered-ua}.
We first build the first saturated graph of local causality
based on the first objective, $\w_1 = \PHobj{a_0}{a_1}$, in the initial context $\ctx$.
This graph is depicted in \pref{fig:metazoan-sa}(left).
%and given \pref{tmh:approxinf}, we have:
We have especially:
\[\allprocs(\thisB{\w_1}{\ctx}) = \{ a_0, c_0, f_1 \} \enspace \text{and}\]
\[\lastprocs(\PHobj{a_0}{a_1}) = \{\{ a_1, c_0, f_1 \}\} \enspace.\]
Therefore, the second step consists of computing the saturated graph of local causality
of the objective $\w_2 = \PHobj{a_1}{a_0}$
in the context
\[\ctx' = \ctx \Cap \allprocs(\thisB{\w_1}{\ctx}) \Cap \lastprocs(\PHobj{a_0}{a_1}) = \PHstate{a_1, c_0, f_1} \enspace.\]
This second graph is given in \pref{fig:metazoan-sa}(middle).
Finally, because
\[\allprocs(\thisB{\w_2}{\ctx'}) = \{ a_1, c_0, c_1, f_1 \} \enspace \text{and}\]
\[\lastprocs(\PHobj{a_1}{a_0}) = \{\{ a_0, c_1 \}\} \enspace,\]
The last step is then checking
the objective $\w_3 = \PHobj{a_0}{a_1}$
in the context
\[\ctx'' = \ctx' \Cap \allprocs(\thisB{\w_2}{\ctx'}) \Cap \lastprocs(\PHobj{a_1}{a_0}) = \PHstate{a_0, c_1, f_1} \enspace.\]
This last graph is depicted in \pref{fig:metazoan-sa}(right).
In conclusion, by using \pref{thm:ordered-ua} on these saturated graphs of local causality,
we can conclude that the initial objective $\w$ is reachable,
and thus that the AAN of \pref{fig:metazoan-ph} is functional.
This result can be extended to the model of \pref{fig:metazoan-php}
because they have the same dynamics (given \pref{th:bisimPHP}).

>>>>

\medskip

At this point, the static analysis results presented in
\pref{ssec:ua} can be used to check if the model is functional,
\ie if gene $a$ can oscillate, thus leading to the formation of stripes.
Starting from context $\ctx^1 = \PHstate{f_1, a_0, c_0}$,
we thus want to check the reachability of $\myp^1 = a_1$;
then, starting from any new state obtained,
we want to check the reachability of $\myp^2 = a_0$,
and finally $\myp^3 = a_1$ once again to ensure that we entered a cycle.
This consists in using \pref{th:approxinf} several times
to check these reachabilities successively.

We first build the local causality graph $\thisB{\myp^1}{\ctx^1}$
related to the reachability of $\myp^1 = a_1$ from the initial state $\ctx^1$.
This graph is depicted in \pref{fig:metazoan-sa}(left).
\pref{th:approxinf} allows to conclude that this reachability is true.
One can thus extract a concretizing scenario from this local causality graph
with a traversal of the graph, as explained in \pref{ssec:concret}.
Such a traversal is detailed in \pref{tab:concret-metazoan};
the scenario extracted from this example is:
$\emptyseq \concat \emptyseq \concat \PHfrappes{c_0, f_1}{a_0}{a_1} =
\PHfrappes{c_0, f_1}{a_0}{a_1}$.
Another traversal would consist in visiting node $f_1$ before node $c_0$
when first descending from node $\{ c_0, f_1 \}$;
however, the same scenario would be extracted.
As there exists no other traversal, we thus have in this case:
$\Delta(\thisB{\myp^1}{\ctx^1}) = \{ \PHfrappes{c_0, f_1}{a_0}{a_1} \}$.

\newcommand{\xproc}[1]{#1}
%\newcommand{\xsol}[1]{\circ^{{#1}}}
\newcommand{\xsol}{\circ}

\begin{table}[htp]
  \centering
  \begin{tabular}{|c|c|c|l|l|l|}
    \hline
    \# & Dir. & Node & Marking & Output & State \\\hline\hline
    1  & $\searrow$ & $\xproc{a_1}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    2  & $\searrow$ & $\PHobj{a_0}{a_1}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    3  & $\searrow$ & $\xsol$ & $\PHfrappes{c_0, f_1}{a_0}{a_1}$ &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    4  & $\searrow$ & $\{ c_0, f_1 \}$ & $\{ c_0, f_1 \}$ &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    \phantom{*}%
    5* & $\searrow$ & $\xproc{c_0}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    6  & $\searrow$ & $\PHobj{c_0}{c_0}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    7  & $\searrow$ & $\xsol$ & $\emptyseq$ & $\emptyseq$ & $\PHstate{f_1, a_0, c_0}$ \\\hline
    8  & $\nearrow$ & $\PHobj{c_0}{c_0}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    9  & $\nearrow$ & $\xproc{c_0}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    10 & $\nearrow$ & $\{ c_0, f_1 \}$ & $\{ f_1 \}$ &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
%    \phantom{*}%
    11 & $\searrow$ & $\xproc{f_1}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    12 & $\searrow$ & $\PHobj{f_1}{f_1}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    13 & $\searrow$ & $\xsol$ & $\emptyseq$ & $\emptyseq$ & $\PHstate{f_1, a_0, c_0}$ \\\hline
    14 & $\nearrow$ & $\PHobj{f_1}{f_1}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    15 & $\nearrow$ & $\xproc{f_1}$ &  &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    16 & $\nearrow$ & $\{ c_0, f_1 \}$ & $\emptyset$ &  & $\PHstate{f_1, a_0, c_0}$ \\\hline
    17 & $\nearrow$ & $\xsol$ & $\emptyseq$ & $\PHfrappes{c_0, f_1}{a_0}{a_1}$ & $\PHstate{f_1, a_1, c_0}$ \\\hline
    18 & $\nearrow$ & $\PHobj{a_0}{a_1}$ &  &  & $\PHstate{f_1, a_1, c_0}$ \\\hline
    19 & $\nearrow$ & $\xproc{a_1}$ &  &  & $\PHstate{f_1, a_1, c_0}$ \\\hline
  \end{tabular}
  \caption{\label{tab:concret-metazoan}%
    Example of the extraction of a concretizing scenario
    from the local causality graph of \pref{fig:metazoan-sa}(left),
    by using the algorithm of \pref{ssec:concret}.
    The first column denotes the step number in the traversal,
    the second depicts the direction of traversal,
    either “$\searrow$” for descending or “$\nearrow$” for ascending,
    the third one is the name of the node,
    the fourth one gives the marking when the node is left,
    the fifth one gives the actions output by the algorithm
    and the last column gives the current state of the model when leaving each node.
    In step \#5, marked with an asterisk,
    the traversal visits node $c_0$, but this was arbitrarily chosen
    amongst the set $\{ c_0, f_1 \}$.
    Another traversal thus consists in visiting node $f_1$ first,
    although in this example it does not change the result.
  }
\end{table}

We denote in the following:
$\ctx^2 = \PHstate{f_1, c_0, a_0} \play \PHfrappes{c_0, f_1}{a_0}{a_1} =
\PHstate{f_1, c_0, a_1}$
the state resulting from the play of a scenario concretizing
$\myp^1$ in $\ctx^1$.
As explained in \todo{ref succ-reach},
One can use this resulting state in order to carry on with another
successive reachability,
such as the reachability of $\myp^2 = a_0$ in this case.
The local causality graph used to check the reachability of $\myp^2$
from $ctx^2$ is given in \pref{fig:metazoan-sa}(middle).
The same reasoning allows to conclude that this reachability is true, and
to extract the following set of concretizing scenarios:
$\Delta(\thisB{\myp^2}{\ctx^2}) =
\{ \PHfrappem{a_1, f_1}{c_0}{c_1} \concat \PHfrappem{c_1}{a_1}{a_0} \}$.
We note that once again, two traversals are possible
(by visiting node $a_1$ before or after note $f_1$)
but both output the same scenario, and therefore end up in the same sate
$\ctx^3 = \PHstate{f_1, a_0, c_1}$.
Finally, the last local causality graph $\thisB{\myp^3}{\ctx^3}$,
depicted in \pref{fig:metazoan-sa}(right),
allows to conclude that this final reachability is true.

In conclusion, by using \pref{th:approxinf} on these local causality graphs,
and the traversal of \pref{ssec:concret} to determine
the starting states of the following reachability step,
we showed that it is possible to reach successively
$a_1$, $a_0$ and $a_1$
and thus that the AAN of \pref{fig:metazoan-ph} is functional.
This result can be extended to the model of \pref{fig:metazoan-php}
because they have the same dynamics (given \pref{th:bisimPHP}).

### 2015-02-13 parts/sa.tex
detailed small example with scenario extraction

A classical approach for determining if a local state of an automaton can be
reached from a given initial state is to build sequences of transitions from the initial state until a state
containing the given local state is reached.
This is essentially what model-checkers do, and the complexity of such analysis (PSPACE-complete
\cite{Harel02}) makes it intractable on large systems, even with advanced symbolic approaches
\cite{PMR12-MSCS}.

Our approach relies on abstractions of scenarios (\pref{ssec:abstr-sce}) that have been introduced in
\cite{PMR12-MSCS} for the static analysis of reachability in the Process Hitting framework,
a particular sub-class of AAN (see the Remark at the end of the previous section).
In this paper, we generalize the static analysis for the case of the under-approximation of
reachability in any AAN (\pref{ssec:ua}).

In \pref{ssec:concret}, we detail a procedure to extract a witness scenario
concretizing a given local state reachability property.
Finally, we discuss in \pref{ssec:ordered-ua} how the static analysis can be
extended to sequential (sub)states reachability properties.

@@

An \emph{objective} (\pref{def:obj}) denotes the reachability of a local state (e.g., $a_j$) of a given automaton $a$
from the initial local state of that automaton (e.g., $a_i$).
Such an objective is written $\PHobjp{a}{i}{j}$.
Successive objectives are described with objective sequences.

@@

  We define an \emph{objective sequence} as a sequence of objectives in which each objective target must be equal to the previous objective bounce of the same automaton, if it exists.
  The set of all objective sequences is denoted by $\OS$.
  Given $\w\in\OS$, $\PHsort(\w) \DEF \{\PHsort(\w_i)\mid i\in\indexes{\w}\}$.
For each automaton $a\in\PHsort(\w)$,  the first local state of $a$ referenced in
$\w$ is denoted by $\first_a(\w) \DEF \target{\w_m}$, where $m=\min\{n\in\indexes{\w}\mid \PHsort(\w)=a\}$.
  The set of objective sequences starting in a state $s\in\PHl$ are denoted by
  $\OS(s)\DEF \{\w\in\OS\mid\forall a\in\PHsort(\w), \first_a(\w)\in s\}$.
\end{definition}

We define the partial ordering relation $\ltw$ between two objective sequences (\pref{def:ltw}) as
follows:
$w\ltw w'$ if and only if there exists a mapping between all the objective bounces of $w'$ in $w$
that preserve the sequentiality, provided that each automaton starts in the same
local state.

\begin{definition}[$\ltw \subset \OS\times\OS$]\label{def:ltw}
$\w\ltw\w'$ if and only if
$|\w|\geq|\w'|$,
$\forall a\in\PHsort(\w'), a\in\PHsort(\w) \wedge \first_a(\w')=\first_a(\w)$;
and there exists a mapping $\phi: \indexes{\w'}\mapsto \indexes{\w}$ such that
$\forall n,m \in\indexes{\w'}, n<m \Leftrightarrow \phi(n)<\phi(m)$,
and
$\forall n\in\indexes{\w'}, \bounce{\w'_n}=\bounce{\w_{\phi(n)}}$.
\end{definition}

@@

An objective sequence can be seen as an abstract representation of a set of scenarios that describe
(part of) the successive state changes of the automata.
We denote by $\concr(\w)$ (\pref{def:concr}) the set of scenarios matching with an objective sequence
$\w$ in the state $s\in\PHs$.
It is essentially all the scenarios for which there exists a mapping from the bounces of each
objective to the bounce of an action in the scenario which preserve the sequential ordering.

\begin{definition}[$\concr: \OS \to \powerset(\Sce)$]\label{def:concr}
Given a state $s\in\PHl$ and an objective sequence
$\w\in\OS(s)$, $\concr(\w)$ is the set of scenarios matching with $\w$:
\begin{align*}
\concr(\w) \DEF \{ \delta\in\Sce(s)\mid &
\w^\vartriangle\neq \emptyseq\Rightarrow
\bounce{\delta_{\card\delta}} = \bounce{\w_{\card\w}}
\\ &
 \wedge \exists \phi:\indexes{\w}\mapsto\indexes{\delta},
    (\forall n,m\in\indexes{\w}, n<m \Leftrightarrow \phi(n)\leq\phi(m))
\\ & \qquad
	\wedge \forall n\in\indexes{\w},
	  \bounce{\w_n} \in s\play\delta_{1..\phi(n)})
\}
\enspace,
\end{align*}
where $\omega^\vartriangle$ refers to the objective sequence $\omega$ where
trivial objectives have been removed.
The notation $\delta_{a..b}$ in the previous definition
denotes the subsequence of $\delta$ between indexes $a$ and $b$,
as defined on page \pageref{notations}.
\end{definition}

@@

Given a non-empty set $\Delta\subseteq\Sce(s)$ of non-empty scenarios that have a common last bounce
($\exists a_i\in\Proc, \forall\delta\in\Delta,
\bounce{\delta_{\card\delta}}=a_i$), one can define an abstraction $\alpha_s$ of
such a set as the smallest objective sequence $\w\in\OS(s)$ such that
$\bounce{\w_{\card\w}} = \bounce{\delta_{\card\delta}}, \delta\in\Delta$
and
$\forall\delta\in\Delta, \exists \phi:\indexes{\w}\mapsto\indexes{\delta}
    (\forall n,m\in\indexes{\w}, n<m \Leftrightarrow \phi(n)\leq\phi(m))
	\wedge \forall n\in\indexes{\w},
	  \bounce{\w_n} \in s\play\delta_{1..\phi(n)})$.
In such a setting, $\alpha_s$ and $\concr$ form a Galois connection between sets
of scenarios and objective sequences.

@@

\begin{example}
Given the AAN of \pref{fig:ph-livelock},
$\BS(\obj{c_0}{c_1}) = \{\PHfrappes{a_1,b_1}{c_0}{c_1}\}$;
and 
$\BS(\obj{a_0}{a_1}) = \{\PHfrappes{b_0}{a_0}{a_1}\}$.
\end{example}

From \pref{def:bs}, we can derive that any scenario matching with an objective includes all the
actions of one of the bounce sequences of the objective (\pref{lem:bs-concr}).
\begin{lemma}
Given a state $s\in \PHl$ and an objective $\obj{a_i}{a_j}$ with $\get{s}{a}=a_i$,
$\forall\delta\in\concr(\obj{a_i}{a_j})$,
$\exists\zeta\in\BS(\obj{a_i}{a_j})$ such that
$\exists \phi: \indexes{\zeta}\to\indexes{\delta}$
with
$\forall n,m\in\indexes{\zeta}, n<m \Leftrightarrow \phi(n)<\phi(m)$
and
$\forall n\in\indexes{\zeta}$, $\zeta_n = \delta_{\phi(n)}$.
\label{lem:bs-concr}
\end{lemma}

@@

Given one initial objective, one could then iteratively refine the objective into objective
sequences following its bounce sequences:
the refinement of objective $P$ with $\zeta\in\BS(P)$ would consists in prepending to $P$ the
objectives leading to the hitters of the actions in $\zeta$, in the same sequential order;
the generalization of the refinement to an objective sequence would require to consider all possible
interleaving between the refinements \cite{PMR12-MSCS}.

For example, given the AAN of \pref{fig:ph-livelock},
in the state $\langle a_0,b_0,c_0\rangle$,
the objective $\obj{a_0}{a_1}$ can be refined in the objective sequence
$\obj{b_0}{b_0}\concat\obj{a_0}{a_1}$
and
the objective $\obj{c_0}{c_1}$ can be refined in the objective sequence
$\obj{a_0}{a_1}\concat\obj{b_0}{b_1}\concat\obj{c_0}{c_1}$
or
$\obj{b_0}{b_1}\concat\obj{a_0}{a_1}\concat\obj{c_0}{c_1}$
(which can then be further refined).

These refinements can be summarized with a so-called Local Causality Graph (LCG).
Given an initial context $\ctx$ and a local state $\myp\in\Proc$,
an LCG $\cwB = (\cwV, \cwE)$
is a digraph where $\cwV \subseteq \Proc\cup \Obj \cup \Sol \cup \sSol$ is the set of vertices
where $\sSol=\powerset(\Proc)$ and $\Sol=\Obj\times\powerset(\sSol)$
and $\cwE \subseteq \cwV \times \cwV$ is the set of oriented edges
(\pref{def:lcg}).
A node in $\cwV\cap\Obj$ represents an objective to refine.
Such a node is related to nodes in $\cwV\cap\Sol$ which represent sets of hitter set of a
bounce sequence of the objective (one for each bounce sequence in $\BS(P)$):
given an objective $P$, if $\zeta\in\BS(P)$, $\abstr\zeta \DEF \{\PHhitter(\zeta_n) \mid
n\in\indexes\zeta \}$, then $P$ is related to $\langle P, \abstr\zeta\rangle\in\Sol$
(\pref{def:lcg}-\ref{lcg-obj-sol}).
For each hitter set $ps\in\abstr\zeta$,
the node $\langle P,\abstr\zeta\rangle\in\cwV\cap\Sol$ is
related to the node $ps\in\sSol$
(\pref{def:lcg}-\ref{lcg-sol-sync}).
Then, each node $ps\in\sSol$ is related to each local state $a\in\Proc$
(\pref{def:lcg}-\ref{lcg-sync-ls}).
Each local state node $a_i\in\cwV\cap\Proc$,
is related to each objective $\obj{a_j}{a_i}$ where
$a_j$ is either in the initial context $\ctx$, or
is a local state node in the LCG
(\pref{def:lcg}-\ref{lcg-ls-obj}).
Finally, given a node $\obj{a_j}{a_i}\in\cwV\cap\Obj$,
if a local state $a_k\neq a_i$ is in its successors (given by
$\conn_{(\cwBNodes,\cwBEdges)}(\obj{a_j}{a_i})$), then
$\obj{a_j}{a_i}$ is related to $\obj{a_k}{a_i}$ in the LCG
(\pref{def:lcg}-\ref{lcg-conn}).

@@

\todo{the procedure is non-deterministic, make it deterministic? (list all the
scenarios}

@@

\todo{name + prototype of the procedure}

\subsection{Addressing Sequential and Sub-state Reachability}
\label{ssec:ordered-ua}
\label{ssec:simult-ua}

@@

In this section, we briefly discuss how the presented static analysis for local
state reachability properties can be extended to sequential (sub)states
reachability properties.

\paragraph{Sub-state reachability.}
Given an AAN $\PH = (\PHs, \PHl, \PHh)$,
a sub-set of automata local states
$ps = \{ a_i \in\Proc\mid a\in\PHs, a_i\in\PHl_a\}$,
the reachability from $s\in\PHl$ of a state $s'\in\PHl$ such that
$ps\subset s'$ is equivalent to the reachability of the local state
$\reach_1$ from a state $s\uplus\langle\reach_0\rangle$
in the AAN
$\PH' = (\PHs', \PHl', \PHh')$ with:
$\PHs' = \PHs \cup \{ \reach \}$, $\PHl' = \PHl \times \PHl_\reach$,
where $\PHl_\reach = \{ \reach_0, \reach_1 \}$,
and $\PHh' = \PHh \cup \{ \PHhit{ps}{\reach_0}{\reach_1} \}$.

Such analysis was not possible with the Process Hitting framework,
because of the lack of the notion of simultaneity for more than two components.

\paragraph{Sequential reachability.}
Given a sequence of local state to reach (e.g., reach $a_i$, then $b_j$),
one can use the same approach as for sub-state reachability by introducing
a new automaton $\reach$ having $n+1$ local states where $n$ is the size of the
sequence of reachability and $n$ actions making it bounce from $\reach_0$ to
$\reach_n$ with hitters corresponding the successive local states
(e.g., $\PHfrappes{a_i}{\reach_0}{\reach_1}$ and
$\PHfrappes{b_j}{\reach_0}{\reach_1}$).

An alternative approach is to use the extraction of a scenario of the previous
sub-section:
given the initial state, and a first local state reachability property, one
compute a possible scenario witnessing this reachability.
Then the next local state reachability properties is computed from the
state that outcomes from the latter scenario.

>>>>

A classical approach for determining if a local state of an automaton can be reach from a
given initial state is to build sequences of transitions from the initial state until a state
containing the given local state is reached.
This is essentially what model-checkers do, and the complexity of such analysis (PSPACE-complete
\cite{Harel02}) makes it intractable on large systems, even with advanced symbolic approaches
\cite{PMR12-MSCS}.

Our approach relies on abstractions of scenarios that have been introduced in
\cite{PMR12-MSCS} for the static analysis of reachability in the Process Hitting framework,
a particular sub-class of AAN (see the Remark at the end of the previous section).
In this paper, we generalize the static analysis for the case of the under-approximation of
reachability in any AAN.

\todo{outline}

@@

An \emph{objective} (\pref{def:obj}) denotes the reachability of a local state (e.g., $a_j$) of a given automaton $a$
from the initial local state of that automaton (e.g., $a_i$).
Such an objective is written $\PHobjp{a}{i}{j}$.
Successive objectives are described with objective sequences (\pref{def:obj}).

@@

  We define an \emph{objective sequence} as a sequence of objectives in which each objective target must be equal to the previous objective bounce of the same automaton, if it exists.
  The set of all objective sequences is denoted by $\OS$.
\end{definition}

We define the partial ordering relation $\ltw$ between two objective sequences (\pref{def:ltw}) as
follows:
$w\ltw w'$ if and only if there exists a mapping between all the objective bounces of $w'$ in $w$
that preserve the sequentiality.

\todo{support compatibility}

\begin{definition}[$\ltw \subset \OS\times\OS$]\label{def:ltw}
$\w\ltw\w'$ if and only if
$|\w|\geq|\w'|$ and
there exists a mapping $\phi: \indexes{\w'}\mapsto \indexes{\w}$ such that
$\forall n,m \in\indexes{\w'}, n<m \Leftrightarrow \phi(n)<\phi(m)$,
and
$\forall n\in\indexes{\w'}, \bounce{\w'_n}=\bounce{\w_{\phi(n)}}$
\end{definition}

@@

An objective sequence can be seen as an abstract representation of a set of scenarios that describe
(part of) the successive state changes of the automata.
We denote by $\concr(\w)$ (\pref{def:concr}) the set of scenarios concretising an objective sequence
$\w$ in the state $s\in\PHs$.
It is essentially all the scenarios for which there exists a mapping from the bounces of each
objective to the bounce of an action in the scenario which preserve the sequential ordering.

\todo{compatibility of $\w$ with $s$}

\begin{definition}[$\concr: \OS \to \powerset(\Sce)$]\label{def:concr}
Given $\w\in\OS$, $\concr(\w)$ is the set of scenarios concretising $\w$ in the
state $s$:
\begin{align*}
\concr(\w) \DEF \{ \delta\in\Sce(s)\mid & (\w^\vartriangle=\emptyseq \wedge
\delta=\emptyseq)  \vee (
\w^\vartriangle\neq \emptyseq\wedge
\\ &
 \wedge \exists \phi:\indexes{\w}\mapsto\indexes{\delta},
    (\forall n,m\in\indexes{\w}, n<m \Leftrightarrow \phi(n)\leq\phi(m))
\\ & \qquad
	\wedge \forall n\in\indexes{\w},
	  \bounce{\w_n} \in s\play\delta_{1..\phi(n)})
\}
\enspace,
\end{align*}
where $\omega^\vartriangle$ refers to the objective sequence $\omega$ where
trivial objectives have been removed.
The notation $\delta_{a..b}$ in the previous definition
denotes the subsequence of $\delta$ between indexes $a$ and $b$,
as defined on page \pageref{notations}.
\end{definition}

@@

@@

From \pref{def:bs}, we can derive that any scenario concretising an objective includes all the
actions of one of the bounce sequences of the objective (\pref{lem:bs-concr}).
\begin{lemma}
Given a state $s\in \PHl$ and an objective $\obj{a_i}{a_j}$ with $\get{s}{a}=a_i$,
$\forall\delta\in\concr(\obj{a_i}{a_j})$,
$\exists\zeta\in\BS(\obj{a_i}{a_j})$ such that
$\exists \phi: \indexes{\zeta}\to\indexes{\delta}$
with
$\forall n,m\in\indexes{\zeta}, n<m \Leftrightarrow \phi(n)<\phi(m)$
and
$\forall n\in\indexes{\zeta}$, $\zeta_n = \delta_{\phi(n)}$.
\label{lem:bs-concr}
\end{lemma}

@@

Given one initial objective, one could then recursively refine the objective into objective
sequences following its bounce sequences \cite{PMR12-MSCS}:
the refinement of objective $P$ with $\zeta\in\BS(P)$ would consists in prepending to $P$ the
objectives leading to the hitters of the actions in $\zeta$, in the same sequential order;
the generalization of the refinement to an objective sequence would require to consider all possible
interleaving between the refinements.

These refinements can be summarized compactly with a so-called Local Causality Graph (LCG).
Given an initial context $\ctx$ and a local state $\myp\in\Proc$,
an LCG $\cwB = (\cwV, \cwE)$
is a digraph where $\cwV \subseteq \Proc\cup \Obj \cup \Sol \cup \sSol$ is the set of vertices
where $\sSol=\powerset(\Proc)$ and $\Sol=\Obj\times\powerset(\sSol)$
and $\cwE \subseteq \cwV \times \cwV$ is the set of oriented edges
(\pref{def:lcg}).
A node in $\Obj$ represents an objective that have to be refined.
Such a node can then be related to nodes in $\Sol$ which represents a set of hitter set of a
bounce sequence of the objective:
given an objective $P$, if $\zeta\in\BS(P)$, $\abstr\zeta \DEF \{\PHhitter(\zeta_n) \mid
n\in\indexes\zeta \}$, then $P$ is related to $\langle P, \abstr\zeta\rangle\in\Sol$
(\pref{def:lcg}-\ref{lcg-obj-sol}).
Given a node $\langle P,\abstr\zeta\rangle\in\cwV\cap\Sol$, for each $ps\in\abstr\zeta$,
it is related to a node $ps\in\sSol$
(\pref{def:lcg}-\ref{lcg-sol-sync}).
A node $ps\in\sSol$ is then related to each $a\in\Proc$
(\pref{def:lcg}-\ref{lcg-sync-ls}).
Given a node $a_i\in\cwV\cap\Proc$, for each $a_j\in\ctx$ or $a_j\in\cwV\cap\Proc$,
it is related to an objective $\obj{a_j}{a_i}$
(\pref{def:lcg}-\ref{lcg-ls-obj}).
Finally, given a node $\obj{a_j}{a_i}\in\cwV\cap\Obj$,
if a local state $a_k\neq a_i$ is in its successors (given by
$\conn_{(\cwBNodes,\cwBEdges)}(\obj{a_j}{a_i})$), then
$\obj{a_j}{a_i}$ is related to $\obj{a_k}{a_i}$ in the LCG
(\pref{def:lcg}-\ref{lcg-conn}).

@@

@@

The execution of such a traversal outputs a scenario.
We denote $\Delta(\cwB)$ the set of all scenarios extracted from the
local causality graph $\cwB$.
Indeed, several scenarios may exist for a same graph,
generated by the successors that can be arbitrarily chosen
in this algorithm.

\subsection{Sequential Under-approximation}
\label{ssec:ordered-ua}

In this section, we briefly explain an alternative sufficient condition that
can be applied when looking for scenarios that successively reach a sequence of local states (e.g.,
first reach $a_j$, then later $b_k$, \ldots).

The approach proposed in this sub-section essentially applies the previous under-approximation for
the first local states, and then computes a set of contexts that may be considered as the starting
point for the next local states.
In that way, the local states to reach are taken into account separately, which implies that only a
subset of scenarios can be captured.
%However, because each iteration focuses on a smaller part of the network, this
%sequential under-approximation may be more conclusive.

\todo{computes possible outcomes from the LCG, gives it as context for the next stage}.

\subsection{Reachability of a State}
\label{ssec:simult-ua}

@@

The reachability property studied so far concerns a single local state at a time.
However, we remark that the reachability of a global state or a sub-state can be
addressed with the very same analysis by introducing a dedicated automata.
Such analysis was not possible with the Process Hitting framework,
because of the lack of the notion of simultaneity for more than two components.

In order to check the reachability of a global state $s \in \PHl$,
let $\PH = (\PHs, \PHl, \PHh)$ be an AAN.
We define a new AAN $\PH' = (\PHs', \PHl', \PHh')$ with:
$\PHs' = \PHs \cup \{ \reach \}$, $\PHl' = \PHl \times \PHl_\reach$,
where $\PHl_\reach = \{ \reach_0, \reach_1 \}$,
and $\PHh' = \PHh \cup \{ \PHhit{\toset{s}}{\reach_0}{\reach_1} \}$.
Given an initial context $\ctx$, the reachability of $s$ in $\PH$
is equivalent to the concretization of $\PHobjp{\reach}{0}{1}$ in $\PH'$
from the initial context $\ctx \cup \{ \reach_0 \}$,
which can be efficiently under-approximated using \pref{th:approxinf}.

It is of course also possible to compute the reachability
of a sub-state $s \in \PHsubl[\PHl]_S$ of a set of components $S \subseteq \PHs$
with the same method.
One can also check the reachability of a set of states $\Lambda \subseteq \PHl$
by creating several actions
$\PHhit{\toset{s}}{\reach_0}{\reach_1}$ in $\PHh$ for each state $s \in \Lambda$.
